---
title: "White-Britt-ADA-data-reanalysis-assignment"
author: "BA White"
date: "4/15/2020"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data reanalysis of: Mahler, D. L., S. M. Lambert, A. J. Geneva, J. Ng, S. B. Hedges, J. B. Losos, and R. E. Glor. 2016. Discovery of a Giant Chameleon-Like Lizard (<i>Anolis</i>) on Hispaniola and Its Significance to Understanding Replicated Adaptive Radiations. The American Naturalist 188:357–364.

#Ecomorphological Affinities
##Goal: Quantify the ecomorphological similarity of new species to other anoles
###Using pPCA Revell (2009) on a covariance matrix of log-transformed and averaged data for body size and 
###shape residuals with the Bayesian MCC chronogram they calculated independent "size" and "shape" axes. 
###load necessary packages for data tidying and transformation
```{r}
library(tidyverse)
library(phytools) 
```
###load log-transformed and averaged size and shape data renamed as a tibble
```{r}
traits <- read_csv("Traits.csv", col_names = TRUE)
```
###see the variables in the tibble
```{r}
glimpse(traits)
```
###The Dryad "Traits.csv" contains 29 more species than their data table results and 2 species are swapped so...
```{r}
traits_3 <- read_csv("Traits_3.csv", col_names = TRUE)
glimpse(traits_3)
```
###these traits do not match what they ultimately included, they excluded 'AVG_lnSVL2' from the analyses.
```{r}
traits <- select(traits_3, c(1:11,13))
glimpse(traits)
```
###load NEXUS tree file - they specify this NEXUS tree in their README file 
```{r}
tree <- read.nexus("5Loc_25milburn_100kthin.MCC.median.nex")
```
###The number of taxa in the data file and tree do not match
```{r}
tip_label <-tibble(tree$tip.label)

traits4tree <- traits %>% 
  semi_join(tip_label, c("species" = "tree$tip.label"))
```
#Trait datadrame must be a MATRIX with species as rownames of that matrix; as required by the pPCA
```{r}
sp <- traits4tree$species
t <- select(traits4tree,-species)
t <- as.matrix(t)
rownames(t) <- sp
```
###prune the branches of the tree for species not found in the morphology dataset 
```{r}
pruned.tree<-drop.tip(tree,tree$tip.label[-match(traits4tree$species, tree$tip.label)])
```
###the 'traits4tree' data have been averaged and log transformed by the original author but they still require size correction between the species pPCA. To do this we will conduct a phylogenetic regression against SVL (using phytools) then collect the residuals for each trait value. 
```{r}
SVL4pgls <- select(traits4tree, AVG_SVL)
SVL4pgls <- as.matrix(SVL4pgls)
rownames(SVL4pgls) <- sp

traits_as_resids <- phyl.resid(pruned.tree, SVL4pgls, t, method="BM")

output <- traits_as_resids$resid
```
###build the model of phylogenetic relationships to the covariance matrix of traits to determine if 
### covariance occurs at a rate greater than expected by Brownian Motion
```{r}
pPCA <- phyl.pca(pruned.tree, output, method="BM", mode="cov")
pPCA
summary(pPCA)
pPCA$S
```
###visualize in screeplot the PCAs
```{r}
plot(pPCA)
```
##Calculate the Euclidean distance to every other species in by the first four axes of a phylogenetic PCA on traits. 
```{r}
scores <- as_tibble(pPCA$S)
fourscores <- select(scores, c(1:4))
fourscores <- as.matrix(fourscores)
rownames(fourscores) <- sp
distance.matrix <- dist(fourscores, method = "euclidean", diag = T) 
distance.matrix <- as.matrix(distance.matrix)
distance.matrix <- as.tibble(distance.matrix)
landestoyialone <- select(distance.matrix, c('landestoyi'))
landestoyialone <- as.matrix(landestoyialone)
rownames(landestoyialone) <- sp
hist(distance.matrix$landestoyi, breaks = 16, right = TRUE)

#The hist from Mahler et al. using dist data from supplements has a mean +0.5 greater than the hist I generated
mahler <- read_csv("euclideanAll.csv", col_names = TRUE)
hist(mahler$Distance,breaks = 16, right = TRUE)

```
##Conducted a 7-category linear discriminant analysis (LDA) on traditional ecomorph class for the four cryptic
##Maybe my histogram is shifted to the left commpard to the Mahler et al 2016 figure because the histogram is actually values genereated after LDA? 
##load ecomorph designation in.
```{r}
ecomo <- select(mahler, c('Species','Ecomorph'))
ecomo1 <- cbind(sp, scores)
ecomo1 <- select(ecomo1, c(1:5)) 
eco_scores1 <- full_join(ecomo, ecomo1, c('Species' = 'sp'))
```
# Linear Discriminant Analysis with Jacknifed Prediction
```{r}
library(MASS)
eco.lda <- lda(Ecomorph ~ PC1 + PC2 + PC3 + PC4, data=eco_scores1)
eco.lda
eco.lda.values <- predict(eco.lda)
##ldahist(data = eco.lda.values$x[,1], g=eco.lda.values$class)
##This histogram works but I don't know the groups.
hist(eco.lda.values$x[,1])
```

```